---
title: "Lesson4 ProblemSet"
output: html_document
---
# Problemset

## Setting up the Environment
```{r}
library('ggplot2')
data(diamonds)
summary(diamonds)
```

## Assignment 1: Plotting a scatterplot of price vs. 'x'

```{r, echo=FALSE}
ggplot(aes(x=x, y=price), data=diamonds) +
  geom_point(color='orange', alpha=.05) +
  coord_trans(y='sqrt') +
  scale_x_discret(breaks = seq(2.5, 10, .5)) +
  xlab('X')

```

## Assignment 2: Correlations
```{r}
with(diamonds, cor(x, price))
with(diamonds, cor(y, price))
with(diamonds, cor(z, price))
```

## Assignment 3: Scatterplot of price vs. depth
```{r}
ggplot(aes(x=depth, y=price), data=diamonds) +
  geom_point(color='orange', alpha=.01) +
  scale_x_continuous(breaks = seq(50, 80, 2))
```

## Assignment 4: Correlation
```{r}
with(diamonds, cor(depth, price))
```

## Assignment 5: Scatterplot of price vs. carat (ommiting top 1% prices and carats)
```{r}
d2 <- subset(diamonds, quantile(price, .99) & quantile(carat, .99))
ggplot(aes(x=carat, y=price), data=d2) +
  geom_point(color='orange')
```

## Assignment 6: Scatterplot of price vs. volume (rough estimate of x*y*z)
```{r}
diamonds$volume <- diamonds$x*diamonds$y*diamonds$z

ggplot(aes(x=volume, y=price), data=diamonds) +
  geom_point()

library(plyr)
count(diamonds$volume == 0)
detach('package:plyr', unload=TRUE)
with(subset(diamonds, diamonds$volume != 0 & diamonds$volume <= 800), cor(volume, price))
diamonds.exclusions <- subset(diamonds, diamonds$volume != 0 & diamonds$volume <= 800)
```

```{r}
ggplot(aes(x=volume, y=price), data=diamonds.exclusions) +
  geom_point(color='orange', alpha=.05) +
  geom_smooth(method='lm')
```

## Assignment 7: Reforming data, plotting clarity and price
```{r}
library(dplyr)
diamondsByClarity <- diamonds %.%
  group_by(clarity) %.%
  summarize(
    mean_price = mean(price),
    median_price = median(price),
    min_price = min(price),
    max_price = max(price),
    n = n()) %.%
  arrange(clarity)

```

## Assignment 8: Mean price by color vs. mean price by clarity
```{r}
library(ggplot2)
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
diamonds_mp_by_clarity.ordered <- summarise(data.frame(factor(diamonds_by_clarity, levels = c("IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2", "I1"))), mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x=clarity, y=mean_price), data=diamonds_mp_by_clarity.ordered) +
  geom_bar(stat="identity")

p2 <- ggplot(aes(x=color, y=mean_price), data=diamonds_mp_by_color) +
  geom_bar(stat="identity")

library(gridExtra)
grid.arrange(p1, p2, ncol=1)
```