Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes: Tracking two variables at the same time. Also there is a neat fit line
running down the center, the closer to which point are, the more accurate the 
guess to the actual number.

```{r Scatterplots}
library(ggplot2)

pf<- read.csv("pseudo_facebook.tsv", sep='\t')

qplot(x=age, y = friend_count, data=pf)

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point()

ggsave('scatter.png')
```

***

#### What are some things that you notice right away?
Response: The dots appear in vertical lines on the page, this appears to be a facet of the nature of finite vs. continuous ages. There are a lot of lies about being > 100.

***

### ggplot Syntax
Notes: ggplot allows for more complicated graphs, but consequently relies on more formal and complex syntax.

```{r ggplot Syntax}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point() +
  xlim(13, 90) # users under 13 not permitted on facebook, over 90 not confident truthful

ggsave('ggplot.png')

summary(pf$age) # gives upper and lower bounds for actual ages 
```

***

### Overplotting
Notes: Overplotting is the phenomenon where it's difficult to tell what the dots are doing because there are so many concentrated in an area.

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20) + #alpha says 'takes 20 points to make one dark point'
  xlim(13, 90)

ggsave('overplot.png')

plot <- ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha=1/20) + # jitter allows correction for finite ages in this case
  xlim(13, 90)

ggsave('overplot_jitter.png')
```

#### What do you notice in the plot?
Response: Lower scores all around, but primarily you see the lack of weight of the outliers. Younger users have much few friends than originally displayed.

***

### Coord_trans()
http://docs.ggplot2.org/current/coord_trans.html
Notes: The difference between transforming the scales and
 transforming the coordinate system is that scale
 transformation occurs BEFORE statistics, and coordinate
 transformation afterwards.  Coordinate transformation also
 changes the shape of geoms:

```{r Coord_trans()}
# syntax: coord_trans(y="sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
plot + ylim(0,5000) + coord_trans(y="sqrt")

# - or - #
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20, position=position_jitter(h=0)) + #positions will still jitter, but minimum height (h) can be no less than 0
  xlim(13, 90) +
  coord_trans(y="sqrt")

ggsave('coord_trans.png')
```

#### What do you notice?
Response: The older end of the scale appeared to rise above the x axis, showing maybe fewer friends close to zero, and the younger end appeared to be more in proportion. The skew brought the numbers lower, and made better use of the space. 

***

### Alpha and Jitter
Notes: Jitter adds noise to points, theoretically pushing values like '0' into the negative, sqrt's of negative numbers are evaluated as imaginary, and unusable for this plot. Geom_point is a better choice, with a modification:

```{r Alpha and Jitter}
ggplot(aes(x=age, y=friendships_initiated), data=pf) +
  geom_point(alpha=1/20, position=position_jitter(h=0)) + 
  xlim(13, 90) +
  coord_trans(y="sqrt")
  
ggsave('fr_init.png')
summary(pf$age)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes: 
install.packages("dplyr")


```{r Conditional Means}

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

